The perceptual dimensions of \\ sonority-driven epenthesis

### 2013-10-21 paper.tex
Edits based on MK's comments

\ex. {\sc Ident(Son$\measuredangle$)}$<$n $\gg$ $\mathcal{M}$
     \a. Scenario 1: {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2}), {\sc Son$\measuredangle$}(C\textsubscript{2},C\textsubscript{3})$<$n \\
         Since neither violates {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n & $\mathcal{M}$ \\[0.5ex]
\hline \hline a. & \ding{43} & C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $<$n & \\
\hline b. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $<$n & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}
      \b. Scenario 2: {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2})$>$n, {\sc Son$\measuredangle$}(C\textsubscript{2},C\textsubscript{3})$<$n \\
          C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} harmonically bound by C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} candidate.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline c. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}  & $<$n & \cellcolor{lightgray} \\
\hline d. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$! & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em} \newpage
     \c. Scenario 3: {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2})$<$n, {\sc Son$\measuredangle$}(C\textsubscript{2},C\textsubscript{3})$>$n \\
         {\sc Ident(Son$\measuredangle$)}$<$n determines output, overruling $\mathcal{M}$.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline e. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$!  & \cellcolor{lightgray} \\
\hline f. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3}& $<$n  & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}
     \d. Scenario 4: {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2}), {\sc Son$\measuredangle$}(C\textsubscript{2}C\textsubscript{3}) $>$ n \\
         Since both candidates violate {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline g. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$  & \\
\hline h. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}

@@

\ex. Scenario: {\sc Son$\measuredangle$}(C\textsubscript{2}C\textsubscript{3}) $>$ {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2}) $>$ n \\

\vspace{-3em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Ident(Son$\measuredangle$)}$<$1.5 &$ \mathcal{M}$  \\[0.5ex]
\hline \hline a. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} &  2.0 $\ast$!  & \cellcolor{lightgray} 2.0 $\ast$ & \cellcolor{lightgray} \\
\hline b.  & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} &  1.7  & \cellcolor{lightgray} 1.7 $\ast$ & \cellcolor{lightgray} $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

Under a universal ranking of {\sc Ident(Son$\measuredangle$)}$<$n constraints, [C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}] would win in this scenario. In an equivalent scenario under free ranking, where {\sc Ident(Son$\measuredangle$)}$<$1.5 is the least stringent constraint to outrank $\mathcal{M}$, which in turn outranks all more stringent {\sc Ident(Son$\measuredangle$)}$<$n constraints, the [C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}] candidate would win instead. There would be conflation of all clusters with {\sc Sonority Angles} below $n$, but distinctions between clusters with {\sc Sonority Angles} would still be relevant.

@@

However, out of the three clusters with a sonority distance of -1, RN has the largest {\sc Sonority Rise}, and therefore is expected to be more resistant to epenthesis than NF and FT. This is contrary to criterion \ref{desideratum_b}. If RN undergoes epenthesis, therefore, we predict FT and NF to do the same.

@@

\ex. \a. More sonorous plateaus are more likely to undergo epenthesis under the {\sc Sonority Angle} metric, whereas all the level clusters should have equal likelihood of being broken up by epenthesis.
     \b. RN is more likely to undergo epenthesis than NF under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RN is more likely to undergo epenthesis than FT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. NF is more likely to undergo epenthesis than FT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RF is more likely to undergo epenthesis than RT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RF is more likely to undergo epenthesis than NT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RT and NT are equally likely to undergo epenthesis under {\sc Sonority Angle}, but NT is more likely to undergo epenthesis than RF under {\sc Sonority Rise}.

@@

Epenthesis does not break up onset sonorant-initial clusters.  This is likely due to the fact that
with a small number of lexical exceptions, Irish words are stress-initial \citep[26]{o.siadhail.1989}.\footnote{With the exception of Munster Irish, where stress is attracted to syllables containing
a long vowel or diphthong \citep{green.1996},
but where onset clusters still are not broken up by
epenthesis.  In this case, one would probably have to posit a positional faithfulness constraint
protecting the word-initial cluster.}  
Breaking up the onset cluster would result in an unstressable epenthetic vowel occupying the 
preferred location of stress in Irish.

@@

    \node (axisstart) at (0.5 * 15,-1.5) {};
    \node (axisend)   at (2.6 * 15,-1.5) {};
    \draw (axisstart) -- (axisend);
   \node (xaxislabel) at (2.6 * 15,-2.0) {\textsc{Sonority Rise}};

@@

\ex. Table showing epenthesis patterns for clusters of varying sonority patterns:

@@

\newpage
\ex. {\sc *CCC} $\gg$ {\sc Dep}

@@

\ex. \a. Behaviour of clusters with C\textsubscript{2}=/r/ and C\textsubscript{3}=sonorant under idiolect \ref{defaultCCCidiolect}:
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-krm-o}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & & \textipa{j@kr1mo} & 1.89 $\ast$! &  \\
\hline b. & \ding{43} & \textipa{j@k1rmo} & 0.12 & $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}		 
\vspace{0.5em}
     \b. Behaviour of clusters with C\textsubscript{2}=/r/ and C\textsubscript{3}=sonorant under idiolect \ref{CRNidiolect}:
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-krm-o}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline c. & \ding{43} & \textipa{j@krimo} & 1.89 & \\
\hline d. & & \textipa{j@kirmo} & 0.12 & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\newpage
\ex. Current ranking predicts incorrect form for cluster /rmd/.

\vspace{-2em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-rmd-o}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \frownie & \textipa{j@rmido} & 2.36 $\ast$! & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{j@rimdo} & 1.89 & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\vspace{-2em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-rmd-o}/} & {\sc Jussive-R} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@rmido} & & \cellcolor{lightgray}2.36 $\ast$ & \cellcolor{lightgray} \\
\hline b. & & \textipa{j@rimdo} & $\ast$! & \cellcolor{lightgray}1.89 & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\vspace{-1em}
\begin{longtable}{ccccc}
	& Underlying form      & Surface form & {\sc Son$\measuredangle$}(C\textsubscript{1},C\textsubscript{2}) & {\sc Son$\measuredangle$}(C\textsubscript{2},C\textsubscript{3}) \\ \hline
    \multicolumn{5}{c}{Fall-\{Rise, Plateau\}: [CC\textipa{1}C]} \\ \hline
	a. & \textipa{j@-rk't'-o} & \textipa{j@nk'1t'o} & 2.36 & 1.37 \\
	b. & \textipa{j@-rks-o}   & \textipa{j@nk1so}   & 2.36 & 0.59 \\
    c. & \textipa{j@-rk'm-o}  & \textipa{j@nk'1mo}  & 2.36 & 0.27 \\
    d. & \textipa{j@-\underline{wzf}-o} & \textipa{j@wz1fo} & 2.03 & 1.33 \\
	e. & \textipa{j@-mxr-o}   & \textipa{j@mx1ro}   & 2.03 & 0.22 \\
    f. & \textipa{j@-\underline{sgd}-o} & \textipa{j@sg1do}   & 2.11 & 1.37 \\  
    g. & \textipa{j@-sdB-o}   & \textipa{j@sd1Bo}   & 2.11 & 0.18 \\  \newpage \hline 
    \multicolumn{5}{c}{\{Rise, Plateau\}-\{Rise, Plateau\}: [C\textipa{1}CC]} \\ \hline
    h. & \textipa{j@-\underline{gdf}-o} & \textipa{j@g1dfo} & 1.37 & 0.59 \\
    i. & \textipa{j@-\underline{gdr}-o} & \textipa{j@g1dro} & 1.37 & 0.12 \\
    j. & \textipa{j@-\underline{kmr}-o} & \textipa{j@k1mro} & 0.27 & 0.46 \\
    k. & \textipa{j@-\underline{sfr}-o} & \textipa{j@s1fro} & 1.33 & 0.22 \\
    l. & \textipa{j@-\underline{sBr}-o} & \textipa{j@s1Bro} & 0.34 & 0.73 \\ \hline
    \multicolumn{5}{c}{Fall-Fall: [CC\textipa{1}C]} \\ \hline
    m. & \textipa{j@-rfk-o} & \textipa{j@nf1ko} & 2.03 & 2.11 \\ 
    n. & \textipa{j@-rmd-o} & \textipa{j@rm1do} & 1.89 & 2.36 \\ 
    o. & \textipa{j-a-mst-o}& \textipa{jams1to} & 2.03 & 2.11 \\ \hline
    \multicolumn{5}{c}{\{Rise,Plateau\}-Fall: [C\textipa{1}CC]} \\ \hline
    p. & \textipa{j@-kft-o} & \textipa{j@k1fto} & 0.59 & 2.11 \\
    q. & \textipa{j@-dmd-o} & \textipa{j@d1mdo} & 0.27 & 2.36 \\
    r. & \textipa{j@-drs-o} & \textipa{j@d1rso} & 0.12 & 2.21 \\ 
    s. & \textipa{j@-\underline{k'rm}-o} & \textipa{j@k'1rmo} & 0.12 & 1.89 \\ 
    t. & \textipa{j@-\underline{srB}-o}  & \textipa{j@s1rBo}  & 0.22 & 1.57 \\
    u. & \textipa{j@-frt-o} & \textipa{j@f1rto} & 0.22 & 2.36 \\
    v. & \textipa{j@-sBx-o} & \textipa{j@s1Bxo} & 0.34 & 2.17 \\
    w. & \textipa{j-a-mrg-o} & \textipa{jam1rgo} & 0.46 & 2.36 \\ \hline
\end{longtable}

@@

\ex. {\sc Align-Left} favours [CC\textipa{1}CC], but {\sc Anchor-Left Rt} favours [C\textipa{1}CC\textipa{1}C].
\vspace{-1em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{n-\textnormal{CCC}-n@}/} & {\sc Align-Left} & {\sc Anchor-Left Rt} \\[0.5ex]
\hline \hline a. & & \textipa{n1\textnormal{C.C}1\textnormal{C}.n@} & 2,3,5,6 & $\ast$! \\
\hline b. & \ding{43} & \textipa{n1.\textnormal{C}1\textnormal{C.C}1.n@} & 2,4!,5,7 & \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

Several possibilities are possible: retain the sonority of the underlying C\textsubscript{1} for that of the output value; use the sonority of the output C\textsubscript{1} as the underlying value; or take the angle formed by the meeting of the two contours, wherever that occurs.

>>>>

\ex. {\sc Ident(Son$\measuredangle$)}$<$n $\gg$ $\mathcal{M}$
     \a. Scenario 1: {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2}), {\sc Son$\measuredangle$}(C\textsubscript{2},C\textsubscript{3})$<$n \\
         Since neither violates {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n & $\mathcal{M}$ \\[0.5ex]
\hline \hline a. & \ding{43} & C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $<$n & \\
\hline b. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $<$n & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}
      \b. Scenario 2: {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2})$>$n, {\sc Son$\measuredangle$}(C\textsubscript{2},C\textsubscript{3})$<$n \\
          C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} harmonically bound by C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} candidate.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline c. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}  & $<$n & \cellcolor{lightgray} \\
\hline d. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$! & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em} \newpage
     \c. Scenario 3: {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2})$<$n, {\sc Son$\measuredangle$}(C\textsubscript{2},C\textsubscript{3})$>$n \\
         {\sc Ident(Son$\measuredangle$)}$<$n determines output, overruling $\mathcal{M}$.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline e. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$!  & \cellcolor{lightgray} \\
\hline f. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3}& $<$n  & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}
     \d. Scenario 4: {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2}), {\sc Son$\measuredangle$}(C\textsubscript{2}C\textsubscript{3}) $>$ n \\ \label{relevantscenario}
         Since both candidates violate {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline g. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$  & \\
\hline h. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}

@@

\ex. Scenario: {\sc Son$\measuredangle$}(C\textsubscript{2}C\textsubscript{3}) $>$ {\sc Son$\measuredangle$}(C\textsubscript{1}C\textsubscript{2}) $>$ n \\ \label{tableauabove}

\vspace{-3em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Ident(Son$\measuredangle$)}$<$1.5 &$ \mathcal{M}$  \\[0.5ex]
\hline \hline a. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} &  2.0 $\ast$!  & \cellcolor{lightgray} 2.0 $\ast$ & \cellcolor{lightgray} \\
\hline b.  &\ding{43}  & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} &  1.7  & \cellcolor{lightgray} 1.7 $\ast$ & \cellcolor{lightgray} $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

As the tableau in \ref{tableauabove} illustrates, under a universal ranking of {\sc Ident(Son$\measuredangle$)}$<$n constraints, [C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3}] would be the winning candidate.

Under free-ranking where {\sc Ident(Son$\measuredangle$)}$<$1.5 is the least stringent constraint to outrank markedness, we have the opposite result.

\ex. Opposite result under free ranking where all more stringent {\sc Ident(Son$\measuredangle$)}$<$n constraints are ranked below $\mathcal{M}$.

\vspace{-2em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} &  {\sc Ident(Son$\measuredangle$)}$<$1.5 &$ \mathcal{M}$ & {\sc Ident(Son$\measuredangle$)}$<$1.9 \\[0.5ex]
\hline \hline a. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} &  2.0 $\ast$ & & \cellcolor{lightgray} 2.0 $\ast$ \\
\hline b. &  & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} &  1.7 $\ast$ &  $\ast$! & \cellcolor{lightgray}  1.7 $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

If both epenthesis sites would result in a {\sc Sonority Angle} larger than the least stringent $n$ to outrank markedness, the candidate that satisfies markedness, [C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}], wins instead. There would be conflation of all clusters with {\sc Sonority Angles} above and below $n$.

@@

However, out of the three clusters FT, NF and RN with a sonority distance of -1, RN has the largest {\sc Sonority Rise}, and therefore is expected to be more resistant to epenthesis than NF and FT. This is contrary to criterion \ref{desideratum_b}. If RN undergoes epenthesis, therefore, we predict FT and NF to do the same by the {\sc Sonority Rise} metric.

@@

\ex. \a. More sonorous plateaus are more likely to undergo epenthesis under the {\sc Sonority Angle} metric, whereas all the level clusters should have equal likelihood of being broken up by epenthesis under {\sc Sonority Rise}.
     \b. RN is more likely to undergo epenthesis than NF under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RN is more likely to undergo epenthesis than FT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. NF is more likely to undergo epenthesis than FT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RF is more likely to undergo epenthesis than RT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RF is more likely to undergo epenthesis than NT under {\sc Sonority Angle}, and vice versa under {\sc Sonority Rise}.
     \b. RT and NT are equally likely to undergo epenthesis under {\sc Sonority Angle}, but NT is more likely to undergo epenthesis than RF under {\sc Sonority Rise}.

@@

Epenthesis does not break up word-initial sonorant-C clusters.  This is likely due to the fact that
with a small number of lexical exceptions, Irish words are stress-initial \citep[26]{o.siadhail.1989}.\footnote{With the exception of Munster Irish, where stress is attracted to syllables containing
a long vowel or diphthong \citep{green.1996},
but where onset clusters still are not broken up by
epenthesis.  In this case, one would probably have to posit a positional faithfulness constraint
protecting the word-initial cluster.}  
Breaking up the onset cluster would result in an unstressable epenthetic vowel occupying the 
preferred location of stress in Irish.

@@

    \node (axisstart) at (0.5 * 15,-1.5) {};
    \node (axisend)   at (2.6 * 15,-1.5) {};
    \draw (axisstart) -- (axisend);
   \node (xaxislabel) at (2.6 * 15,-2.0) {\textsc{SonRise}};

@@

\ex. Table showing epenthesis patterns for clusters of varying sonority patterns:\footnote{No specific examples are available in \citep{rose.2000} for the gaps in the table, but their pattern of epenthesis is implied in the text and Rose's analysis.}

@@

\ex. {\sc *CCC} $\gg$ {\sc Dep}

@@

\ex. \a. Behaviour of clusters with C\textsubscript{2}=/r/ and C\textsubscript{3}=sonorant under idiolect \ref{defaultCCCidiolect}:
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-krm-o}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & & \textipa{j@kr1mo} & 1.89 $\ast$! &  \\
\hline b. & \ding{43} & \textipa{j@k1rmo} & 0.12 & $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}		 
\vspace{0.5em}
     \b. Behaviour of clusters with C\textsubscript{2}=/r/ and C\textsubscript{3}=sonorant under idiolect \ref{CRNidiolect}:
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-krm-o}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline c. & \ding{43} & \textipa{j@kr1mo} & 1.89 & \\
\hline d. & & \textipa{j@k1rmo} & 0.12 & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. Current ranking predicts incorrect form for cluster /rmd/.

\vspace{-2em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-rmd-o}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \frownie & \textipa{j@rm1do} & 2.36 $\ast$! & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{j@r1mdo} & 1.89 & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\vspace{-2em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-rmd-o}/} & {\sc Jussive-R} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@rm1do} & & \cellcolor{lightgray}2.36 $\ast$ & \cellcolor{lightgray} \\
\hline b. & & \textipa{j@r1mdo} & $\ast$! & \cellcolor{lightgray}1.89 & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\vspace{-1em}
\begin{longtable}{ccccc}
	& Underlying form      & Surface form & {\sc Son$\measuredangle$}(C\textsubscript{1},C\textsubscript{2}) & {\sc Son$\measuredangle$}(C\textsubscript{2},C\textsubscript{3}) \\ \hline
    \multicolumn{5}{c}{Fall-\{Rise, Plateau\}: [CC\textipa{1}C]} \\ \hline
	a. & \textipa{j@-rk't'-o} & \textipa{j@nk'1t'o} & 2.36 & 1.37 \\
	b. & \textipa{j@-rks-o}   & \textipa{j@nk1so}   & 2.36 & 0.59 \\
    c. & \textipa{j@-rk'm-o}  & \textipa{j@nk'1mo}  & 2.36 & 0.27 \\
    d. & \textipa{j@-\underline{wzf}-o} & \textipa{j@wz1fo} & 2.03 & 1.33 \\
	e. & \textipa{j@-mxr-o}   & \textipa{j@mx1ro}   & 2.03 & 0.22 \\
    f. & \textipa{j@-\underline{sgd}-o} & \textipa{j@sg1do}   & 2.11 & 1.37 \\  
    g. & \textipa{j@-sdB-o}   & \textipa{j@sd1Bo}   & 2.11 & 0.18 \\  \hline 
    \multicolumn{5}{c}{\{Rise, Plateau\}-\{Rise, Plateau\}: [C\textipa{1}CC]} \\ \hline
    h. & \textipa{j@-\underline{gdf}-o} & \textipa{j@g1dfo} & 1.37 & 0.59 \\
    i. & \textipa{j@-\underline{gdr}-o} & \textipa{j@g1dro} & 1.37 & 0.12 \\
    j. & \textipa{j@-\underline{kmr}-o} & \textipa{j@k1mro} & 0.27 & 0.46 \\
    k. & \textipa{j@-\underline{sfr}-o} & \textipa{j@s1fro} & 1.33 & 0.22 \\
    l. & \textipa{j@-\underline{sBr}-o} & \textipa{j@s1Bro} & 0.34 & 0.73 \\ \hline
    \multicolumn{5}{c}{Fall-Fall: [CC\textipa{1}C]} \\ \hline
    m. & \textipa{j@-rfk-o} & \textipa{j@nf1ko} & 2.03 & 2.11 \\ 
    n. & \textipa{j@-rmd-o} & \textipa{j@rm1do} & 1.89 & 2.36 \\ 
    o. & \textipa{j-a-mst-o}& \textipa{jams1to} & 2.03 & 2.11 \\ \hline
    \multicolumn{5}{c}{\{Rise,Plateau\}-Fall: [C\textipa{1}CC]} \\ \hline
    p. & \textipa{j@-kft-o} & \textipa{j@k1fto} & 0.59 & 2.11 \\
    q. & \textipa{j@-dmd-o} & \textipa{j@d1mdo} & 0.27 & 2.36 \\
    r. & \textipa{j@-drs-o} & \textipa{j@d1rso} & 0.12 & 2.21 \\ 
    s. & \textipa{j@-\underline{k'rm}-o} & \textipa{j@k'1rmo} & 0.12 & 1.89 \\ 
    t. & \textipa{j@-\underline{srB}-o}  & \textipa{j@s1rBo}  & 0.22 & 1.57 \\
    u. & \textipa{j@-frt-o} & \textipa{j@f1rto} & 0.22 & 2.36 \\
    v. & \textipa{j@-sBx-o} & \textipa{j@s1Bxo} & 0.34 & 2.17 \\
    w. & \textipa{j-a-mrg-o} & \textipa{jam1rgo} & 0.46 & 2.36 \\ \hline
\end{longtable}

@@

\ex. {\sc Align-Left} favours [CC\textipa{1}CC], but {\sc Anchor-Left Rt} favours [C\textipa{1}CC\textipa{1}C].
\vspace{-1em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{n-\textnormal{CCC}-n@}/} & {\sc Align-Left} & {\sc Anchor-Left Rt} \\[0.5ex]
\hline \hline a. & & \textipa{n1\textnormal{C.C}1\textnormal{C}.n@} & 2,3,5,6 & $\ast$! \\
\hline b. & & \textipa{n1.\textnormal{C}1\textnormal{C.C}1.n@} & 2,4!,5,7 & \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

Several solutions are possible: retain the sonority of the underlying C\textsubscript{1} for that of the output value; use the sonority of the output C\textsubscript{1} as the underlying value; or take the angle formed by the meeting of the two contours, wherever that occurs.
